---
title: "Debugging"
author: "Yayehirad A Melsew"
date: "02/06/2021"
output: html_document
---
when we run the software, it does something different from what we expected. In these situations, we need to engage in a process to determine if

Our expectations were incorrect, based on the documented behavior of the software; or
There is a problem with the code, such that the programming is not done in a way that will match expectations.
This is the process of debugging.

R comes with a set of built-in tools for interactive debugging that can be useful for tracking down the source of problems. These functions are

browser(): an interactive debugging environment that allows you to step through code one expression at a time
debug() / debugonce(): a function that initiates the browser within a function
trace(): this function allows you to temporarily insert pieces of code into other functions to modify their behavior
recover(): a function for navigating the function call stack after a function has thrown an error
traceback(): prints out the function call stack after an error occurs; does nothing if there’s no error

# traceback()
If an error occurs, the easiest thing to do is to immediately call the traceback() function. This function returns the function call stack just before the error occurred so that you can see what level of function calls the error occurred. If you have many functions calling each other in succeeding, the traceback()output can be useful for identifying where to go digging first.
```{r}
check_n_value <- function(n) {
        if(n > 0) {
                stop("n should be <= 0")
        }
}
error_if_n_is_greater_than_zero <- function(n){
        check_n_value(n)
        n
}
error_if_n_is_greater_than_zero(5)
#Error in check_n_value(n): n should be <= 0
traceback()
```
From the traceback, we can see that the error occurred in the check_n_value() function. Put another way, the stop() function was called from within the check_n_value() function.

# Browsing a Function Environment
```{r}
check_n_value <- function(n) {
        if(n > 0) {
                browser()  ## Error occurs around here
                stop("n should be <= 0")
        }
}

error_if_n_is_greater_than_zero(5)

 browser()
```

# Tracing Functions
If you have easy access to the source code of a function (and can modify the code), then it’s usually easiest to insert browser() calls directly into the code as you track down various bugs. However, if you do not have easy access to a function’s code, or perhaps a function is inside a package that would require rebuilding after each edit, it is sometimes easier to make use of the trace() function to make temporary code modifications.
```{r}
check_n_value <- function(n) {
        if(n > 0) {
                
                stop("n should be <= 0")
        }
}

error_if_n_is_greater_than_zero(5)

 trace()
```
#Using debug() and debugonce()
The debug() and debugonce() functions can be called on other functions to turn on the “debugging state” of a function. Calling debug() on a function makes it such that when that function is called, you immediately enter a browser and can step through the code one expression at a time.
The debugging state is persistent, so once a function is flagged for debugging, it will remain flagged. Because it is easy to forget about the debugging state of a function, the debugonce() function turns on the debugging state the next time the function is called, but then turns it off after the browser is exited.
```{r}
## Turn on debugging state for 'lm' function
debug(lm)
undebug(lm)
```

# recover()
The recover() function is not often used but can be an essential tool when debugging complex code. Typically, you do not call recover() directly, but rather set it as the function to invoke anytime an error occurs in code. This can be done via the options()function.

```{r}
options(error = recover)
error_if_n_is_greater_than_zero(5)
```
Usually, when an error occurs in code, the code stops execution and you are brought back to the usual R console prompt. However, when recover() is in use and an error occurs, you are given the function call stack and a menu.
The recover() function is very useful if an error is deep inside a nested series of function calls and it is difficult to pinpoint exactly where an error is occurring (so that you might use browser() ortrace()).
Another scenario is when there is a stochastic element to your code so that errors occur in an unpredictable way. Using recover() will allow you to browse the function environment only when the error eventually does occur.
# Summary
Debugging in R is facilitated with the functions browser,debug, trace, recover, and traceback.
These debugging tools should not be used as a crutch when developing functions.




