---
title: "Functional Programming with purr package"
author: "Yayehirad A Melsew"
date: "30/05/2021"
output: html_document
---
Functional programming concentrates on four constructs - 
- Data (numbers, strings, etc), 
- Variables (function arguments),
- Functions, and 
- Function Applications (evaluating functions given arguments and/or data).
There are groups of functions that are essential for functional programming. In most cases they take a function and a data
| structure as arguments, and that function is applied to that data structure in some way. The purrr library contains many of
| these functions and we’ll be using it throughout this lesson. Functional programming is concerned mostly with lists and
| vectors.
few functions that we are going to use throughout this lesson

# Map functions
Mapping is a powerful technique for thinking about how to apply computational operations to your data.
The map family of functions applies a function to the elements of a data structure, usually a list or a vector. The function
| is evaluated once for each element of the vector with the vector element as the first argument to the function. The return
| value is the same kind if data structure (a list or vector) but with every element replaced by the result of the function
| being evaluated with the corresponding element as the argument to the function.
In the purrr package the map() function returns a list, while the map_lgl(), map_chr(), and map_dbl() functions return vectors
| of logical values, strings, or numbers respectively.

```{r}
library(purrr)

map_chr(c(5, 4, 3, 2, 1), function(x){
  c("one", "two", "three", "four", "five")[x]
})
#[1] "five"  "four"  "three" "two"   "one"  

map_lgl(c(1, 2, 3, 4, 5), function(x){
  x > 3
})

#or
map_lgl(c(1, 2, 3, 4, 5), gt, b = 3)
#[1] FALSE FALSE FALSE  TRUE  TRUE
```
 A predicate function is a function that returns TRUE or FALSE for each element in the provided list or vector
 The map_if() function takes as its arguments a list or vector containing data, a predicate function, and then a function to be applied. A predicate function is a function that returns TRUE orFALSE for each element in the provided list or vector. In the case ofmap_if(): if the predicate functions evaluates to TRUE, then the function is applied to the corresponding vector element, however if the predicate function evaluates to FALSE then the function is not applied. The map_if() function always returns a list, so I’m piping the result of map_if() to unlist() so it look prettier:Notice how only the even numbers are squared, while the odd numbers are left alone.
```{r}
map_if(1:5, function(x){
              x %% 2 == 0
            },
            function(y){
              y^2
            }) %>% unlist()
```

The map_at() function only applies the provided function to elements of a vector specified by their indexes. map_at() always returns a list so like before I’m piping the result to unlist():Like we expected to happen the providied function is only applied to the first, third, and fifth element of the vector provided.
```{r}
map_at(seq(100, 500, 100), c(1, 3, 5), function(x){
  x - 10
}) %>% unlist()
```
map a function over two data structures with the map2() family of functions. The first two arguments should be two vectors of the same length, followed by a function which will be evaluated with an element of the first vector as the first argument and an element of the second vector as the second argument. For example:
```{r}
map2_chr(letters, 1:26, paste)
```
The pmap() family of functions is similar to map2(), however instead of mapping across two vectors or lists, you can map across any number of lists. The list argument is a list of lists that the function will map over, followed by the function that will applied:
```{r}
pmap_chr(list(
  list(1, 2, 3),
  list("one", "two", "three"),
  list("uno", "dos", "tres")
), paste)
```

# Reduce family of functions
List or vector reduction iteratively combines the first element of a vector with the second element of a vector, then that
| combined result is combined with the third element of the vector, and so on until the end of the vector is reached. The
| function to be applied should take at least two arguments. Where mapping returns a vector or a list, reducing should return a
| single value.










